### 2020-02-25 paper/revisions/review-responses.tex

        \begin{solution}
                Thank you very much for these comments. We appreciate your detailed review, which has certainly improved the paper. The manuscript has been enhanced and more detail is described regarding these changes in the specific comment responses below.
        \end{solution}

        %---------------------------------------------------------------------

        \question SD-TMSR concept development does not correctly described in the section "introduction", where it states that the SD-TMSR was introduced by the Chinese Academy of Sciences. Actually, SD-TMSR concept was proposed by ORNL as early as in the 1960s, which was called Molten salt breeding reactor (MSBR) (see Ref.: Tech. Rep. ORNL-4541). In addition, the study of reference 7 which was cited to illustrate SD-TMSR concept actually adopted the single-flow single-zone MSR rather than SD-TMSR.

        \begin{solution}
                 Thank you for catching this. Just small correction: this ORNL 
                 concept was called \emph{Molten Salt Breeder Reactor (MSBR)}. 
                 The whole statement has been 
                 modified to read:\\

                 The Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR) with a thermal power of 2,250
                 MW$_{th}$ was proposed for the first time by ORNL as early as 
                 in the 1960s, which was called Molten Salt Breeder Reactor 
                 (MSBR) \cite{robertson_conceptual_1971}.

@@

        \end{solution}

        %---------------------------------------------------------------------

        \question The authors did not conduct a sufficient literature survey on the fuel cycle transition studies for the SD-TMSR concept. Cui. et al., have made an extensive study on fuel cycle transition for SD-TMSR by employing LEU and Pu as the startup fissile materials and two transition modes (in-core transition and ex-core transition) (see Ref.: Transition toward thorium fuel cycle in a molten salt reactor by using plutonium, NUCL SCI TECH (2017) 28:152; Possible scenarios for the transition to thorium fuel cycle in molten salt reactor by using enriched uranium. Progress in Nuclear Energy 104 (2018) 75-84). The state "For the SD-TMSR concept, simulating the fissile core startup and fuel cycle transition with other fissile materials (except U-233) has not studied before" in page 3 is therefore inaccurate.
        \begin{solution}
                Thank you for your helpful review. The statement "For the SD-TMSR 
                concept, simulating the fissile core startup and fuel cycle 
                transition with other fissile materials (except$^{233}$U) has 
                not studied before" in page 3, has been removed. 
                The following paragraphs have been added to the manuscript:\\

                The research effort described in \cite{cui2017transition} and
                \cite{cui2018possible} is most similar to the work
                presented in this paper. Cui \emph{et al.} analyzed the fuel 
                transition from 
                enriched $^{235}$U/Th and Pu/Th to $^{233}$U/Th for the 
                MSBR-like design. Cui 
                \emph{et al.} found that the fuel transition can be achieved 
                by using: (1) 
                enriched uranium with greater than 40\% enrichment 
                \cite{cui2018possible}; (2) 
                Pu from LWR (burn-up of 60 GWd/t) spent nuclear fuel 
                \cite{cui2017transition, 
                	cui2018possible}.

                However, a few major differences follow: 
                (1) Cui \emph{et al.} employed an in-house tool named MSR 
                Reprocessing 
                Sequence (MSR-RS) \cite{ZOU2015114} which couples with SCALE 
                and employs a 
                batch-wise approach while we used built-in truly continuous 
                fuel depletion 
                capabilities in SERPENT-2 \cite{aufiero2013extended}; 
                (2) Cui \emph{et al.} studied a high-enriched uranium ($>$40\% 
                of $^{235}$U) 
                and plutonium partitioned from light water reactors spent fuel 
                as the startup 
                fissile materials while we considered different loading cases: 
                HALEU, Pu 
                mixed with HALEU, reactor-grade plutonium, TRU and $^{233}$U;
                (3) we adopted the SD-TMSR core geometry optimized by Li 
                \emph{et al.} 
                \cite{li_optimization_2018} which is different from the 
                MSBR-like geometry by 
                Mathieu \emph{et al.} \cite{mathieu_possible_2009} adopted by 
                Cui \emph{et 
                	al.}; 
                (4) Cui \emph{et al.} introduced two scenarios for thorium 
                fuel cycle 
                transition: a Breeding and Burning (B\&B) scenario and a 
                Pre-breeding scenario 
                while we adopting different approach by introducing two 
                different feed 
                mechanisms: thorium and non-thorium.

                Comparing with Cui \emph{et al.}, we used different reactor 
                design (geometry, 
                volumes, densities, fuel salt compositions) and computation 
                method (continuous 
                online reprocessing instead of batch-wise approach). In Cui's 
                scenarios, the 
                amount of thorium is kept constant. $^{233}$U is fed only in 
                (B\&B) scenario 
                and extra fissile materials would have to be added into the 
                core if the 
                produced $^{233}$U is not enough to maintain the reactor 
                criticality. In the 
                Pre-breeding scenario, all the $^{233}$U produced from the 
                decay of the 
                extracted $^{233}$Pa stored outside the core until it reaches 
                the required 
                startup mass for a new reactor. Thus, the criticality of the 
                core is 
                maintained by refueling external fissile material. In the 
                current work, using 
                the thorium feed mechanism, we simultaneously feed thorium and 
                all or fraction 
                of produced $^{233}$U from the \texttt{Pa-decay tank} into the 
                core. The 
                excess $^{233}$U (if exists) is stored outside the core. In 
                the non-thorium 
                feed mechanism, we continuously inject external heavy metals 
                (HALEU, Pu mixed 
                with HALEU, reactor-grade Pu, or TRU) and fraction of produced 
                $^{233}$U from 
                the \texttt{Pa-decay tank}. For the non-thorium mechanism, we 
                feed a part of 
                produced $^{233}$U because in some cases (e.g., TRU) refueling 
                only external 
                heavy metals failed to maintain the core criticality 
                (limitation is 
                dm$_{salt}$ $\leq$ 0.1\%).

@@

        \end{solution}

        %---------------------------------------------------------------------

        \question More details should be provided to illustrate the heavy metal addition for the thorium feed mechanism and non-thorium feed mechanism in page 11. It is hard to understand why the mass of Th / external heavy metals (excluding Th) is approximately equal to the mass of extracted Pa.
        \begin{solution}
               Thank you for the comment. Actually, in page 11, we mention 
               that the mass of Th / external heavy metals is approximately 
               equal to the mass of extracted Fission products (FPs) and not 
               Pa. Because in both feed mechanisms (thorium and non-thorium) 
               we continuously added $^{233}$U and the mass of injected 
               $^{233}$U is equal to the mass of extracted Pa ($^{233}$Pa 
               decays to $^{233}$U after $\approx$ 27 d). To keep the total 
               fuel mass almost constant, for \textbf{thorium feed mechanism}, 
               the mass of injected thorium must be equal to the mass of 
               extracted FPs and for \textbf{non-thorium feed mechanism}, the 
               mass of injected fuel (i.e., HALEU, Pu mixed with HALEU, 
               reactor-grade Pu, or TRU) must be equal to the mass of 
               extracted FPs. We rewrite the paragraph on page 11 to clarify 
               this point:\\

               \begin{enumerate}
               	\item The simulation starts without injecting refueling materials (i.e. only removing FPs and Pa).
               	\item After the first depletion calculation step, we check the total mass density of FPs and Pa in the \texttt{FP-waste tank} and \texttt{Pa-decay tank}, respectively.
               	\item A simple calculation yields the amount of heavy metal that must be added during this cycle:
               	\subitem 3a. \textbf{Thorium feed mechanism}: mass of injected 
               	$^{233}$U $\approx$ mass of extracted Pa and mass of injected 
               	$^{232}$Th $\approx$ mass of extracted FPs to maintain the 
               	total fuel mass constant.
               	\subitem 3b. \textbf{Non-thorium feed mechanism}: mass of injected $^{233}$U $\approx$ mass of extracted Pa and mass of injected fuel (i.e., HALEU, Pu mixed with HALEU, reactor-grade Pu, or TRU) $\approx$ mass of extracted FPs to maintain the total fuel mass constant.
               	\item Dividing this mass by time and inventory of refueling material gives the corresponding feed constant (1/s).
               \end{enumerate}
               In both feed mechanisms (thorium and non-thorium) we 
               continuously added $^{233}$U and the mass of injected $^{233}$U 
               is equal to the mass of extracted Pa ($^{233}$Pa decays to 
               $^{233}$U after $\approx$ 27 d). To keep the total fuel mass 
               almost constant, for \textbf{thorium feed mechanism}, the mass 
               of injected thorium must be equal to the mass of extracted FPs 
               and for \textbf{non-thorium feed mechanism}, the mass of 
               injected fuel (i.e., HALEU, Pu mixed with HALEU, reactor-grade 
               Pu, or TRU) must be equal to the mass of extracted FPs.

@@

        \end{solution}

        %--------------------------------------------------------------------

        \question Phenomenons presented in Figure 3 that K-eff for the case of Pu+HALEU firstly increases and then decreases, and the time to become subcritical for the case of reactor-grade Pu is longest (except the case of U-233) should be explained.
        \begin{solution}
                 Thank you for your insightful comment. Phenomenons presented 
                 in Figure 3 have been described in the manuscript.
                 The following paragraph and additional Figure~4 has been added to the manuscript:\\

                 For the Pu+HALEU case, $k_{eff}$ firstly increases during the 
                 first $\approx$ $3$ EFPY and then gradually decreases. This 
                 is due to the conversion of $^{238}$U from HALEU into fissile 
                 plutonium, which often produces the majority of the neutrons 
                 until the transition to Th/$^{233}$U fuel cycle. Figure~4 
                 shows the variation in fissile plutonium isotopes ($^{239}$Pu 
                 , $^{241}$Pu, and $^{243}$Pu) inventory during reactor 
                 operation. For TRU and reactor-grade Pu, the fissile 
                 plutonium isotopes are gradually depleted, during operating 
                 (the fissile material is dominated by Pu in these cases). For 
                 the HALEU case, the fissile material is dominated by 
                 $^{235}$U, 
                 however, $^{238}$U is converted into fissile plutonium. 
                 Therefore, the mass of fissile plutonium increases during 
                 operating but its value still relatively low. For the 
                 Pu+HALEU case, the mass of the fissile plutonium isotopes 
                 increases 
                 within the first $\approx$ $3$ EFPY and then decreases. The 
                 reactivity from the produced Pu is much greater than the 
                 depletion of the initial reactivity (due to the depletion of 
                 initial fissile isotopes), consequently, $k_{eff}$ increases 
                 during the first $\approx$ $3$ years and then decreases; the 
                 amount of $^{233}$U generated in the Pu+HALEU case is 
                 insufficient to maintain the reactor criticality and 
                 counteract neutron absorption in non-fissile isotopes.\\

@@

        \begin{solution}
                 We appreciate your detailed review.                   
                 The following paragraph has been added to the manuscript:\\

                 The neutron spectrum of TRU case at BOL is softer than $^{233}$U case because $^{232}$Th inventory is much lower for the TRU case: $54.4$ $t$ instead of $76.9$ $t$ for the $^{233}$U. $^{232}$Th has a very high absorption cross section in the thermal region ($\approx$ $10^2$ b). Moreover, $^{232}$Th has resonance region between $10^{-5}$ and $10^{-3}$ which justify relatively low energy of neutrons for the TRU case in this energy range (see Figure~14).

        \end{solution}

        %--------------------------------------------------------------------
        \question More details should be provided to explain the radial 
        neutron flux distribution for the cases of Pu reactor-grade and TRU 
        presented in Fig.13 and 14. For both the radial fast and thermal 
        neutron flux change from BOL to EOL, why it is much larger for Pu 
        reactor-grade case compared with TRU case. In addition, the state "In 
        contrast, for the reactor-grade Pu and TRU cases, significant changes 
        are observed for thermal neutron flux in the outer core zone and 
        reactor" is inaccurate, apparently from Figure 14, it can be seen that 
        the Pu reactor-grade (initial) case has a small change in the outer 
        core zone and reflector. 
        \begin{solution}
				Thank you for pointing it out. 
				At the EOL the fuel salt composition is more or less the same 
				(``equilibrium'' composition), hence, both fast and thermal 
				fluxes looks very similar for plutonium and TRU cases.
				At the BOL, this difference is caused by different $^{232}$Th 
				inventory at BOL for Pu reactor-grade (67.5t) and TRU (54.4t). 
				 $^{232}$Th is a strong absorber in the thermal and 
				 intermediate energy range, so, it causes such a significant  
				 difference in the flux: reactor-grade Pu composition provides 
				 much harder spectrum (see Figure~14) at BOL.

                Figure 15 shows 
				significant changes \textbf{during reactor operation} in the 
				outer core zone for the Pu (green line) and almost no changes 
				for the TRU (orange). However, we are agree that thermal flux 
				drop in the reflector is more dramatic for the TRU (initial 
				and equilibrium compositions) and the Pu (equilibrium) cases 
				than for the U-233 case. We think it happens because of 
				non-optimal geometry of the core for the Pu and TRU initial 
				loading.

				The following paragraphs have been added to the manuscript for 
				clarification:\\

@@

				Notably, the neutron flux change for both the fast and thermal 
				energies is much larger for the reactor-grade Pu case than for 
				TRU case (Figures~15 and 16). The reason is the much softer 
				neutron spectrum for the TRU initial loading compared with the 
				reactor-grade Pu as shown in Figure~14: peak thermal neutron 
				flux is about 4 times larger for the TRU case. The neutron 
				spectrum in the SD-TMSR core initially loaded by TRU is much 
				softer than the reactor-grade Pu case and even than the U-233 
				case because $^{232}$Th inventory is the lowest for the TRU 
				case: $54.4$ $t$ instead of $67.5$ $t$ and $76.9$ $t$ for the reactor-grade  
				Pu and U-233 cases, respectively (Table~5). $^{232}$Th greatly 
				affects the neutron energy spectrum due to its large 
				absorption cross section in the thermal energy range and 
				resonance region between $10^{-5}$ and $10^{-3}$. Finally, 
				when the TRU and the reactor-grade Pu cases reached 
				equilibrium fuel salt composition ($\approx 30$ years after 
				startup), the radial fast and thermal neutron flux 
				distribution became similar in both shape and magnitude.

				More changes in thermal neutron flux shape and magnitude 
				during the reactor operation for the $^{233}$U case are 
				observed in the inner core zone ($R\lesssim150$) than in the 
				outer core zone. In the outer core zone, significant changes 
				during reactor operation are observed for the thermal 
				neutron flux for the reactor-grade Pu case. In contrast, for 
				the TRU case, thermal neutron flux change in the outer core 
				zone  is almost negligible. Additionally, 
				Figure~16 shows relatively significant changes in thermal flux 
				leakage from the core for the Pu (equilibrium composition) and 
				TRU (initial and equilibrium composition) cases. The SD-TMSR 
				core design was optimized for $^{233}$U  				
				\cite{li_optimization_2018}; thus, the core geometry (e.g., 
				channel lattice pitch) must be re-optimized for another type 
				of fuel to obtain better neutronics performance.
        \end{solution}

### 2020-02-25 paper/results.tex
Edits after meeting with Huff

Notably, the molar fractions of \gls{HALEU} and reactor-grade Pu in the initial fuel composition were increased
more. Consequently, the initial $k_{eff}$ for these cases increased (Figure~\ref{fig:keff1} (HALEU and reactor-grade Pu cases)).
For the HALEU fuel salt, the amount of $^{233}$U generated in the SD-TMSR after a few years is insufficient to counteract the absorption of the (non-fissile) $^{238}$U added at start-up after much of the original $^{235}$U is depleted. Consequently, the core becomes subcritical after $\approx$ $6$ years of operation.

For Pu+HALEU case, $k_{eff}$ firstly increases during the first $\approx$ $3$ EFPY and then gradually decreases. This is due to the conversion of $^{238}$U from HALEU into fissile plutonium, which often produces the majority of the neutrons until the transition to Th/$^{233}$U fuel cycle. Figure~\ref{fig:diffPuAll811} shows the variation in fissile plutonium isotopes ($^{239}$Pu , $^{241}$Pu, and $^{243}$Pu) inventory during reactor operation. For TRU and reactor-grade Pu, the fissile plutonium isotopes are gradually depleted, during operating (the fissile material is dominated by Pu in these cases). For HALEU case, the fissile material is dominated by $^{235}$U, however, $^{238}$U is converted into fissile plutonium. Therefore, the mass of fissile plutonium increases during operating but its value still relatively low. For Pu+HALEU case, the mass of the fissile plutonium isotopes increases within the first $\approx$ $3$ EFPY and then decreases. The reactivity from the produced Pu is much greater than the depletion of the initial reactivity (due to the depletion of initial fissile isotopes), consequently, $k_{eff}$ increases during the first $\approx$ $3$ years and then decreases; the amount of $^{233}$U generated in the Pu+HALEU case is insufficient to maintain the reactor criticality and counteract neutron absorption in non-fissile isotopes.

@@

More changes in thermal neutron flux shape and magnitude during the reactor 
operation for the $^{233}$U case are observed in the inner core zone 
($R\lesssim150$) than in the outer core zone. In the outer core zone, 
significant changes during reactor operation are observed for the thermal 
neutron flux for the reactor-grade Pu case. In contrast, for the TRU case, it 
is almost negligible. Additionally, Figure~\ref{fig:thermal_flux} shows 
relatively significant changes in thermal flux leakage from the core for the 
Pu (equilibrium composition) and TRU (initial and equilibrium composition) 
cases. The SD-TMSR core design was optimized for $^{233}$U  
\cite{li_optimization_2018}; thus, the core geometry (e.g., channels lattice 
pitch) must be re-optimized for another type of fuel to obtain better 
neutronics performance.

>>>>

Notably, the molar fractions of \gls{HALEU} and reactor-grade Pu in the initial fuel composition were increased
more. Consequently, the initial $k_{eff}$ for these cases increased (see 
Figure~\ref{fig:keff1} for HALEU and reactor-grade Pu cases).
For the HALEU fuel salt, the amount of $^{233}$U generated in the SD-TMSR after a few years is insufficient to counteract the absorption of the (non-fissile) $^{238}$U added at start-up after much of the original $^{235}$U is depleted. Consequently, the core becomes subcritical after $\approx$ $6$ years of operation.

For the Pu+HALEU case, $k_{eff}$ firstly increases during the first $\approx$ 
$3$ EFPY and then gradually decreases. This is due to the conversion of 
$^{238}$U from HALEU into fissile plutonium, which often produces the majority 
of the neutrons until the transition to Th/$^{233}$U fuel cycle. 
Figure~\ref{fig:diffPuAll811} shows the variation in fissile plutonium 
isotopes ($^{239}$Pu , $^{241}$Pu, and $^{243}$Pu) inventory during reactor 
operation. For TRU and reactor-grade Pu, the fissile plutonium isotopes are 
gradually depleted, during operating (the fissile material is dominated by Pu 
in these cases). For the HALEU case, the fissile material is dominated by 
$^{235}$U, however, $^{238}$U is converted into fissile plutonium. Therefore, 
the mass of fissile plutonium increases during operating but its value still 
relatively low. For the Pu+HALEU case, the mass of the fissile plutonium 
isotopes 
increases within the first $\approx$ $3$ EFPY and then decreases. The 
reactivity from the produced Pu is much greater than the depletion of the 
initial reactivity (due to the depletion of initial fissile isotopes), 
consequently, $k_{eff}$ increases during the first $\approx$ $3$ years and 
then decreases; the amount of $^{233}$U generated in the Pu+HALEU case is 
insufficient to maintain the reactor criticality and counteract neutron 
absorption in non-fissile isotopes.

@@

More changes in thermal neutron flux shape and magnitude during the reactor 
operation for the $^{233}$U case are observed in the inner core zone 
($R\lesssim150$) than in the outer core zone. In the outer core zone, 
significant changes during reactor operation are observed for the thermal 
neutron flux for the reactor-grade Pu case. In contrast, for the TRU case, 
thermal neutron flux change in the outer core zone is almost negligible. 
Additionally, Figure~\ref{fig:thermal_flux} shows 
relatively significant changes in thermal flux leakage from the core for the 
Pu (equilibrium composition) and TRU (initial and equilibrium composition) 
cases. The SD-TMSR core design was optimized for $^{233}$U  
\cite{li_optimization_2018}; thus, the core geometry (e.g., channel lattice 
pitch) must be re-optimized for another type of fuel to obtain better 
neutronics performance.
